Seminarium TI 3700\\ Data Analysis on Twitter: An Overview (working title)

### 2014-03-05 authorinstr_2.tex
last changes

\subsection{Twitter API}
\section{Single tweet analysis}
In this section we will discuss the different types of analysis that can be done on one single tweet. This usually is also the basis for a analysis of a larger amount of tweets or even on a stream of tweets. 
\subsection{Data enrichment}
At the very basis of a tweet is data enrichment. Because the length of a tweet is restricted to 140 characters the users of Twitter use language in creative ways that sometimes make it, even for humans, hard to understand what a tweet is about. To get a better understanding of the meaning of a tweet, a context is needed. For humans this is done by linking the information from the tweet to earlier acquired knowledge.
\subsubsection{Data enrichment issues}
The problem of linking tweets to another source of information poses the biggest issue in data enrichment. The tweets themselves do not provide much information. Fortunately, Kwak et al. have shown that the topic of over 85\% of the of the tweet is related to news \cite{newsmedia}. This knowledge can be leveraged to create a context for most of the tweets.  This is what what Abel et al. proposed to do\cite{enrichmentForProfiling}. They linked tweets to mainstream news networks like CNN, BBC and New York Times.

Linking to just news articles narrows the problem significantly, however, actually linking the tweets and the new articles still is an issue. Abel et al. propose two main linking strategies, divided into five methods total.
\begin{enumerate}
\item \textbf{URL-based strategy}\\
This strategy is based upon the observation that many tweets contains URLs. Usually these URLs refer to news articles the tweets are closely related to.
	\subitem \textbf{Strict URL-based} defines an article and a tweet as related when the tweet contains an URL to that news article.
	\subitem \textbf{Lenient URL-based}, this method is an extension of the strict URL-based method. This extension entails that if a tweet $t$ has a link to an article $a$, all tweets that are retweeting $t$ considered linked to $a$. The same counts for tweets that are in responds to $t$, they will also be linked to $n$.
\item \textbf{Content-based strategy}
	\subitem \textbf{Bag of words}, in this method a tweet is represented as a vector of word frequencies. For a news article we can do the same thing. To link an article to a tweet we need to calculate the similarity between the tweet and the article and pick the one with the highest similarity. The similarity is calculated by the $TF \times IDF$ (Term Frequency times Inverse Document Frequency): the number of times a word from the tweet occurs in a news article, multiplied by the logarithm of the total number of articles over the number of articles the word occurs in. 
	\subitem \textbf{Hashtag-based} is based upon the same concept as the Bag of words method. The difference here is that we do not consider the words in the article but we are only interested in the hashtags in the tweet for our representation of the tweet.
	\subitem \textbf{Entity-based} is also based upon the Bag of words method. The entity-based method does not consider words in an article. It relies on the entities that are mentioned in the article to make a vector to represent the article. 
\end{enumerate}
Using these method a broader context can be created. This still is not much more useful than the tweet on its own. For the article to be actually useful the topics of the articles have to be extracted. To do this Natural Language Processing and machine learning can be used to extract the entities from the article. OpenCalais\footnote{www.opencalais.com} is a web service that does just that. It returns not only the entities, it will also return facts and events hidden in the text, which will make the result much more useful.

\subsection{Sentiment analysis}
Sentiment analysis or opinion mining is "a linguistic analysis technique where a body of text is examined to characterize the tonality of the document."\footnote{http://lexicon.ft.com/Term?term=sentiment-analysis} In other words, the goal of a sentiment analysis is to extract the writers opinion on some topic from a body of text. This body of text can very well be a tweet, and that combined with the popularity of microblogging amongst internet users, gives a very interesting body of knowledge. But in order to get to that knowledge the data has to be analyzed fist. 
The most important and most obvious issue is how to extract an opinion from a tweet. In the next part we will discuss a few ways to extract an opinion from a tweet.
\subsubsection{Semantics }
The most important way to extract a sentiment from a tweet is by analyzing the semantics. After all, tweets are written text. The tweets can be labeled using machine learning algorithms. Pang et al. \cite{machineLearning} have done a comparative study of three standard machine learning algorithms: 
\begin{enumerate}
\item \textbf{Naive Bayes} is a fair simple classifier that operates under the assumption that the presents or absence of features does not relate to the presents of other features. If we apply that on Twitter we would get $c = \arg\max_{c\in C} P(c|w)$ where $c$ is a sentiment class, $C$ the collection of sentiment classes and $w$ is a tweet. \cite{sentAnalysis}
\item \textbf{Maximum Entropy} is a model that, unlike Bayes, does not make an independence assumption. It operates under the philosophy that "we should choose the model making the fewest assumptions about the data while still remaining consistent with it".\cite{machineLearning} \\
The formula is of the form $P(c|d) = \frac{1}{Z(d)}\exp(\Sigma_i \lambda_{i,c}F_{i,c}(d,c))$, where $Z(d)$ is a normalization function, $\lambda_{i,c}$ a weighted vector determining the significance of the feature and $F_{i,c}$ is a class/feature function. 
\item \textbf{Support Vector Machine} is fundamentally different from Naive Bayes and Max Entropy because it is not a probabilistic model. The basic idea behind a support vector machine is that a vector is trained to separate the sentiments. Not only will it separate the training data, it will do so at the largest possible margin. 
\end{enumerate}
It turns out that over the whole the Support Vector Machine performs the best in the sentiment analysis\cite{machineLearning}. The differences, however, are not substantially large.

\subsubsection{Emoticons}
Other than semantical analysis we can also use other data from tweets. A good sentiment indicator are emoticons. Twitter users use emoticons (e.g. ":-)", ":(" or "=(") to indicate their mood.\cite{twitterSentiment} Since the number of emoticons is limited and therefor a sentiment can be assigned easily, the information from the emoticons can be leveraged in two ways
\begin{enumerate}
\item \textbf{Labeling:} The emoticons can be used to assign a sentiment to a tweet. This can be a very useful asset to take into account when assigning sentiments. The emoticons can be considered unigrams for some other classifier. There they will be strong evidence for a certain sentiment.
\item \textbf{Training:} Because of their clear evidence for a sentiment emoticons can also be used to build a training corpus. Twitter can be queried for "positive" emoticons and "negative" emoticons and by assuming the emoticon reflects the sentiment of the tweet, it can be labeled accordingly.\cite{moodLens} 
\end{enumerate}

\subsection{Concept extraction}
\subsection{Metadata analysis (opt)}

@@

\subsubsection{User Classification Issues}
\textbf{Context}\\
To classify users on their tweets is it useful to extract the context of the content of a tweet. Some hashtag $x$ could have a different context in two tweets. So extra information about the content of a tweet is needed. For instance \cite{userclasst} uses mentions of (members of) political parties to deduce if a certain hashtag is used in a political context. 
\\\\
\textbf{Profile}\\
Profile information can be helpful to classify a user, but it should be used with care due to the amount of fake, inconsistent or incorrect information. According to \cite{usermachine} 80\% of the users provide a location and 48\% of the users provide a short bio. Furthermore etnicity and gender are far more difficult to extract. 
\subsection{Tag based analysis}
\subsection{Event identification}

>>>>

\subsection{Twitter API}
\section{Single tweet analysis}
In this section we will discuss the different types of analysis that can be done on one single tweet. This usually is also the basis for a analysis of a larger amount of tweets or even on a stream of tweets. 
\subsection{Data enrichment}
At the very basis of the analysis of a tweet is data enrichment. Because the length of a tweet is restricted to 140 characters the users of Twitter use language in creative ways that sometimes make it, even for humans, hard to understand what a tweet is about. To get a better understanding of the meaning of a tweet, a context is needed. For humans this is done by linking the information from the tweet to earlier acquired knowledge. The same thing can be done by computers. 
\subsubsection{Context creation}
The problem of linking tweets to another source of information poses the biggest issue in data enrichment. The tweets themselves do not provide much information. Fortunately, Kwak et al. have shown that the topic of over 85\% of the of the tweets is related to news \cite{newsmedia}. This knowledge can be leveraged to create a context for most of the tweets.  This is what Abel et al. proposed to do\cite{enrichmentForProfiling}. They linked tweets to mainstream news networks like CNN, BBC and New York Times.

Linking to just news articles narrows the problem significantly, however, actually linking the tweets and the new articles still is an issue. Abel et al. propose two main linking strategies, divided into five methods total.
\begin{enumerate}
\item \textbf{URL-based strategy}\\
This strategy is based upon the observation that many tweets contains URLs. Usually these URLs refer to news articles the tweets are closely related to. Two URL-based strategies are proposed
	\subitem \textbf{Strict URL-based} defines an article $a$ and a tweet $t$ to be related when $t$ contains an URL to $a$.
	\subitem \textbf{Lenient URL-based}, this method is an extension of the strict URL-based method. This extension entails that if a tweet $t$ has a link to an article $a$, all tweets that are retweeting $t$ are considered linked to $a$. The same counts for tweets that are in responds to $t$, these tweets will also be considered linked to $a$.
\item \textbf{Content-based strategy}\\
Content-based strategies rely on the idea of finding as much similarity as possible between the content of a tweet and the content of a news article. The most similar tweet-article pair is considered to be related and therefor they are linked.
	\subitem \textbf{Bag of words}, in this method a tweet is represented as a vector of word frequencies. For a news article we can also create a vector the represent it. To link an article to a tweet we need to calculate the similarity between the tweet and the article and pick the one with the highest similarity. The similarity is calculated by the $TF \times IDF$ (Term Frequency times Inverse Document Frequency): the number of times a word from the tweet occurs in a news article, multiplied by the logarithm of the total number of articles over the number of articles the word occurs in. \cite{tfidf}
	\subitem \textbf{Hashtag-based} is based upon the same concept as the Bag of words method. The difference is that we do not consider the words in the article, instead we only use the hashtags in the representation of a tweet to calculate the $TF \times IDF$ with.
	\subitem \textbf{Entity-based} is also based upon the Bag of words method. The entity-based method does not consider words in an article. It relies on the entities that are mentioned in the article to make a vector to represent the article. 
\end{enumerate}
Using these method a broader context can be created. This still is not much more useful than the tweet on its own. For the article to be actually useful the topics of the articles have to be extracted. To do this Natural Language Processing and machine learning can be used to extract the entities from the article. OpenCalais\footnote{www.opencalais.com} is a web service that does just that. It returns not only the entities, it will also return facts and events hidden in the text, which will make the result much more useful.

\subsection{Sentiment analysis}
Sentiment analysis or opinion mining is "a linguistic analysis technique where a body of text is examined to characterize the tonality of the document."\footnote{http://lexicon.ft.com/Term?term=sentiment-analysis} In other words, the goal of a sentiment analysis is to extract the writers opinion on some topic from a body of text. This body of text can very well be a tweet, and that combined with the popularity of microblogging amongst internet users, makes Twitter a very interesting body of knowledge about the opinion of a large group of people. But in order to get to that knowledge the data has to be analyzed fist. 
The most important and most obvious issue is how to extract an opinion from a tweet. In the next part we will discuss a few ways to extract an opinion from a tweet.
\subsubsection{Semantics }
The most important way to extract a sentiment from a tweet is by analyzing its semantics. After all, tweets are written text. A sentiment can be assigned to a tweet using machine learning algorithms. Pang et al. \cite{machineLearning} have done a comparative study of three standard machine learning algorithms for sentiment analysis: 
\begin{enumerate}
\item \textbf{Naive Bayes} is a fair simple classifier that operates under the assumption that the presents or absence of features does not relate to the presents of other features. If we apply that on Twitter we would get $c = \arg\max_{c\in C} P(c|w)$ where $c$ is a sentiment class, $C$ the collection of sentiment classes and $w$ is a tweet. \cite{sentAnalysis}
\item \textbf{Maximum Entropy} is a model that, unlike Bayes, does not make an independence assumption. It operates under the philosophy that "we should choose the model making the fewest assumptions about the data while still remaining consistent with it".\cite{machineLearning} \\
The formula is of the form $P(c|d) = \frac{1}{Z(d)}\exp(\Sigma_i \lambda_{i,c}F_{i,c}(d,c))$, where $Z(d)$ is a normalization function, $\lambda_{i,c}$ a weighted vector determining the significance of the feature and $F_{i,c}$ is a class/feature function. 
\item \textbf{Support Vector Machine} is fundamentally different from Naive Bayes and Max Entropy because it is not a probabilistic model. The basic idea behind a support vector machine is that a vector is trained to separate the sentiments. Not only will it separate the training data, it will do so at the largest possible margin. 
\end{enumerate}
It turns out that over the whole the Support Vector Machine performs the best on the sentiment analysis\cite{machineLearning}. The differences, however, are not substantially large.

\subsubsection{Emoticons}
Other than semantical analysis we can also use other data from tweets. A good sentiment indicator are emoticons. Twitter users use emoticons (e.g. ":-)", ":(" or "=(") to indicate their mood.\cite{twitterSentiment} Since the number of emoticons is limited a sentiment can easily be assigned manually. The information from the emoticons can be leveraged in two ways
\begin{enumerate}
\item \textbf{Labeling:} The emoticons can be used to assign a sentiment to a tweet. This can be a very useful asset to take into account when assigning a sentiment to a tweet. The emoticons can be considered unigrams feature for some classifier. There they will be strong evidence for a certain sentiment.
\item \textbf{Training:} Because of their clear evidence for a sentiment emoticons can also be used to build a training corpus for a supervised learning algorithm. Twitter can be queried for "positive" emoticons and "negative" emoticons and then, by assuming the emoticon reflects the sentiment of message, the tweet can be labeled accordingly.\cite{moodLens} 
\end{enumerate}
We now see that emoticons can be used to extract a sentiment from a tweet. However, they are not sufficient to just use emoticons to label the tweets. Simply because not all tweets contain emoticons.
\subsubsection{Usage}
The possibilities of sentiment analysis are not too hard to imagine. Every company wants to know how their newly launched product is doing, film makes want to know what people think about their new movie and politicians want to know how they are doing for the next elections. Using sentiment analysis all this information will become available\cite{machineLearning}. 
\subsection{Concept extraction}
\subsection{Metadata analysis (opt)}

@@

\subsubsection{User Classification Issues}
\textbf{Context}\\
To classify users on their tweets is it useful to extract the context of the content of a tweet. Some hashtag $x$ could have a different context in two tweets. So extra information about the content of a tweet is needed. For instance \cite{userclasst} uses mentions of (members of) political parties to deduce if a certain hashtag is used in a political context. 
\\\\
\textbf{Profile}\\
Profile information can be helpful to classify a user, but it should be used with care due to the amount of fake, inconsistent or incorrect information. According to \cite{usermachine} 80\% of the users provide a location and 48\% of the users provide a short bio. Furthermore etnicity and gender are far more difficult to extract.

\subsection{Tag based analysis}

Twitter allows their users to contribute their own social annotations to their tweets in the form of tags. These tags can be arbitrarily generated by users 
which make them hold interesting information about both the tweet and the user that posted it.

\subsection{Event identification}
